---
title: "traffic-data"
author: "Mina Sedra"
date: "April 6, 2016"
output: html_document
---
Installing and adding the Libraies we will use
```{r,results="hide", warning=FALSE}
library("dplyr", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library(ggplot2)
library(plyr)
library('tm')
library('wordcloud')
library(corrgram)
```

Loading the data 
```{r, warning=FALSE}
data <- read.csv(file = '/Users/mina/Desktop/traffic-data.csv')
```

Cleaning Data:

first removing the constant columns and the NAs
```{r, warning=FALSE}
same <- sapply(data, function(.col){all(.col[1L] == .col)})
same <-as.vector(same)
same[is.na(same)] <- FALSE
dataset <- data[!same]
```

Storing the classes of the columns

```{r, warning=FALSE}
classes.of.all.columns <- sapply(data, class)
```

Then removing the duplicates  

```{r, warning=FALSE}
duplicated_rows = duplicated(dataset$rd.rp.cmid);
dataset <- dataset[!duplicated_rows,]
```

Now the Size of the data is 

```{r, warning=FALSE}
dim(dataset)
```

Since stid has a meanig
 stid    meaning
  6      question
  7      radar
  8      accident
  9      3otl
  10     attention
  
Then replacing the NAs in it is according to the comments

```{r, warning=FALSE}
question <- is.na(dataset$rd.rp.stid) & (grepl("[?]", dataset$rd.rp.cm, ignore.case=TRUE)|grepl("[؟]", dataset$rd.rp.cm, ignore.case=TRUE))
dataset$rd.rp.stid[question] <- 6


radar <-is.na(dataset$rd.rp.stid) & (grepl("رادار", dataset$rd.rp.cm, ignore.case=TRUE)|grepl("radar", dataset$rd.rp.cm, ignore.case=TRUE))
dataset$rd.rp.stid[radar] <- 7


accident <- is.na(dataset$rd.rp.stid) & (grepl("حادثة", dataset$rd.rp.cm, ignore.case=TRUE)|grepl("حادثه", dataset$rd.rp.cm, ignore.case=TRUE)|grepl("7adsa", dataset$rd.rp.cm, ignore.case=TRUE))
dataset$rd.rp.stid[accident] <- 8


Aotl <- is.na(dataset$rd.rp.stid) & (grepl("3otl", dataset$rd.rp.cm, ignore.case=TRUE)|grepl("عطلانة", dataset$rd.rp.cm, ignore.case=TRUE))
dataset$rd.rp.stid[Aotl] <- 9

#Adding a vector to say is this report is from bey2ollak

from.GPS.Reporter <-dataset$rd.rp.fullnm == "GPS Reporter"
from.GPS.Reporter[is.na(from.GPS.Reporter)] <- FALSE
  

```

Add the boolean vectors to the data 

```{r, warning=FALSE}
dataset["question"] <- question
dataset["radar"] <- radar
dataset["accident"] <- accident
dataset["Aotl"] <- Aotl
dataset["Report"] <- from.GPS.Reporter

head(dataset)
```

Add the time where the report actually issued 
```{r, warning=FALSE}

#change the crawl date postixct
dataset$crawl_date <- as.POSIXct(dataset$crawl_date,format = "%a %b  %d %H:%M:%S",tz = "UTC")

#add reported at 
reported.at <- dataset$crawl_date -(dataset$rd.rp.hr * 60 * 60) - (dataset$rd.rp.mn * 60);
dataset["reported.at"] <- reported.at;

dataset <- dataset %>% mutate(report.range= format(reported.at, format="%y-%m-%d %H"))

```

And Visualizing the Correlations 
```{r, warning=FALSE}
corrgram(dataset, order=NULL, lower.panel=panel.shade,upper.panel=NULL, text.panel=panel.txt,main="coleration")
```

Divide the road name to 3 cluomns {from , to , through}
```{r, warning=FALSE}
splited.str<-t(data.frame(strsplit(as.character(dataset$rd.nm), "[;]")))
dataset<-dataset%>%mutate(through=splited.str[,1],rd.nm=splited.str[,2])

splited.str <- t(data.frame(strsplit(as.character(dataset$rd.nm), "To")))
dataset<-dataset%>%mutate(from=splited.str[,1],to=splited.str[,2])

```

So the dataset now looks like 
```{r, warning=FALSE}
head(dataset)
```

As there is no need in our analysis to have (ad.logo,crawl_date,rd.img)
we will drop them


```{r, warning=FALSE}

drops <- c("rd.img","ad.logo","crawl_date","rd.stid","rd.hr","rd.mn","rd.new","rd.strq","rd.cmrq")
dataset <- dataset[ , !(names(dataset) %in% drops)]
head(dataset)
```




Get the speed from the report
first make new dataset from the GPS Reporters only
Then add the speed 

```{r, warning=FALSE}
data.from.GPS.Reporters <- dataset[from.GPS.Reporter,]

pattern.e <-".*\\[(\\d+ km/h)\\].*"
pattern.a <-".*\\[(\\d+ كم/س)\\].*"

s.e =  gsub(pattern.e, "\\1",dataset$rd.rp.cm[from.GPS.Reporter]) 
s.a =  gsub(pattern.a, "\\1",dataset$rd.rp.cm[from.GPS.Reporter]) 

s.a[nchar(s.a)>10] <- "0"
s.a <- as.numeric(gsub("([0-9]+).*$", "\\1", s.a))
s.a[is.na(s.a)]= 0
s.e[nchar(s.e)>10] <- "0"
s.e <- as.numeric(gsub("([0-9]+).*$", "\\1", s.e))
s.e[is.na(s.e)]= 0

data.from.GPS.Reporters["speed"] <- as.numeric(gsub("([0-9]+).*$", "\\1", s.a)) + as.numeric(gsub("([0-9]+).*$", "\\1", s.e))

head(data.from.GPS.Reporters)

```

To see what the people's most fequent words used to express that the road is 7alawa so we will use
Function to analyze the comments and get the frequency the words in the comments with specific stid
```{r, warning=FALSE}


comment_analysis <- function(n){
rdata <- dataset[dataset$rd.rp.stid == n,  ]
rdata <- rdata[!rdata$Report,]
review_text <- paste(rdata$rd.rp.cm, collapse=" ")

review_source <- VectorSource(review_text)
corpus <- Corpus(review_source)

corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, removeWords, c("إلى","شارع","أخد","الطريق","kmh","كم/س","average","speed","لحد"))

dtm <- DocumentTermMatrix(corpus)
dtm2 <- as.matrix(dtm)

frequency <- colSums(dtm2)
frequency <- sort(frequency, decreasing=TRUE)
words <- names(frequency[1:20])
wordcloud(words[1:20], frequency[1:20])
return (frequency[1:20])
}
```

So for 7alawa

```{r,results="hide", warning=FALSE}
comment_analysis(1);
```

And the same for lazeez
```{r,results="hide", warning=FALSE}
comment_analysis(2)
```


And for mashy
```{r,results="hide", warning=FALSE}
comment_analysis(3)
```

And for za7ma
```{r,results="hide", warning=FALSE}
comment_analysis(4)
```

And for mafeesh 2amal
```{r,results="hide", warning=FALSE}
comment_analysis(5)
```

Also for radars
```{r,results="hide", warning=FALSE}
comment_analysis(7)
```

And for 3otl

```{r,results="hide", warning=FALSE}
comment_analysis(9)


```
 
 
 
 Also we can know the top 5 active users
 
```{r,results="hide", warning=FALSE}
users.count <- table(dataset$rd.rp.fullnm)
users.count <- sort(users.count, decreasing=TRUE)
users.count[1:5]
```

And of course GPS Reporter is the most active user so this histogram only for users 
```{r,results="hide", warning=FALSE}
barplot(users.count[2:5], main="Top Active Users")
```

Now analyse by road so for exapmle we will take "Ghamra To Masged ElNour"

So this is will show a graph the speed vs time for this road and it will color the points according to the report type 
1 - 7alawa
2 - lazeez
3 - mashy
4 - za7ma
5 - mafeesh 2amal

```{r,results="hide", warning=FALSE}
Ghamra_Masged.ElNour = data.from.GPS.Reporters[data.from.GPS.Reporters$rd.nm == "Ghamra To Masged ElNour" ,]
ggplot(aes(x = Ghamra_Masged.ElNour$report.range, y = Ghamra_Masged.ElNour$speed), data = Ghamra_Masged.ElNour ,labels()) +  geom_point(aes(colour = factor(Ghamra_Masged.ElNour$rd.rp.stid))) + labs(x = "Time",y = "Speed",colour = "Report Type")+expand_limits(x=c(0,30))

```

And for all time 

```{r,results="hide", warning=FALSE}
Ghamra_Masged.ElNour = dataset[dataset$rd.nm == "Ghamra To Masged ElNour" ,]

ggplot(Ghamra_Masged.ElNour, aes(x=rd.rp.stid )) +
  geom_histogram(binwidth=.5, alpha=.5, position="identity") + labs(x = "Report type",y = "count",colour = "Report Type")

```

